### 2016-12-19 coverletter/CoverLetter.tex
DT review comments

%\documentclass[ngerman]{ethbrief3}
\documentclass[english]{ethbrief3}
%\documentclass[ngerman,DepLogo]{ethbrief3}
%\documentclass[english,DepLogo]{ethbrief3}
%\documentclass[ngerman,ETHUZH]{ethbrief3}
%\documentclass[english,ETHUZH]{ethbrief3}
% \usepackage{silence}
% \WarningFilter{latex}{Option pagecontinue is false}
\usepackage[utf8]{inputenc}
\usepackage[pdfborder={0 0 0},pdftex]{hyperref}
\usepackage{graphicx}

% automatic language support (Niklas Beisert)
\EthLang{
  \RequirePackage[document]{ragged2e}% Corrects hyphenation in german
  \RequirePackage{ngerman}
  \PassOptionsToPackage{ngerman}{babel}
}{
  \PassOptionsToPackage{english}{babel}
}

% Declaration of the personal data
    \EthFirstUnitName{\EthLang{1. Zeile Organisationseinheit}{Institute of Technology in Architecture}}
    % \EthSecondUnitName{\EthLang{2. Zeile Organisationseinheit}{Future Cities Laboratory}}
    \EthName{ETH \Zurich} % if to change the town see below
    \EthPersName{\EthLang{Herr Max Mustermann}{Mr. Clayton Miller}}
    \EthPersPosition{\EthLang{Funktion}{Researcher}}
    \EthOffice{\EthLang{Geb\"aude Raum}{Future Cities Laboratory}}
    \EthStreet{\EthLang{Strasse Hausnummer}{1 CREATE Way \#06-01}}
    \EthTown{Singapore} % if to change the town see below
%    \EthAddress{\EthLang{Freie\\Adresseingabe}{freestyle\\address}}
% --- country optional, typical if sendig outside CH
%     \EthLang{}{\EthCountry{\Switzerland}} % only the country in english version
     %\EthCountry{Singapore} % in german and english version
     \EthPhone {(Worldwide Contact Number):\\+1 402 403 0090}
%     \EthPhone {+41 44 632 bb bb} % secondary phone, fax, \ldots
%     \EthFax   {+41 44 632 aa ab}
%     \EthMobile{+41 7 59 aa bbb}
\EthLang{
     \EthEmail{miller.clayton@arch.ethz.ch}
     \EthWeb{www.systems.arch.ch}
}{
     \EthEmail{miller.clayton@arch.ethz.ch}
     \EthWeb{www.systems.arch.ch}
}
%     \EthPersInfo{\EthLang{Zus\"atzliche\\Daten}{additional\\data}}

% --- end of declaration of the personal data ---

% --- load the signature-graphic
%\signaturefile[0.5cm]{\includegraphics[height=1cm]{mysig}} % include a signature file
\signature{Clayton Miller}

%\makelabels         % to get the address for an envelope
%\date{July, 3 2013} % be free to define the date manually
% --- multilinguale Ortsnamen, \Zurich definiert als Standard
\location{Singapore}  % be free to define the location manually

\subject[]{Journal Paper Submission}

\recipient{
Ian Beausoleil-Morrison \\
Jan Hensen\\
Co-Editors\\
Journal of Building Performance Simulation\\}

\begin{document}
\opening{Dear Dr. Beausoleil-Morrison and Dr. Hensen:}

The attached manuscript entitled “Urban and building multiscale co-simulation: case study implementations on two university campuses” is an original research article that I would like to submit for consideration in Journal of Building Performance Simulation. This paper outlines the process of co-simulation between EnergyPlus and CitySim for two campuses. This paper builds upon the work from two previous conference papers and includes properly-referenced content from both papers. References of those papers are found below: \\[12pt]
Thomas, D., Miller, C., Kämpf, J., Schlueter, A., 2014. Multiscale co-simulation of EnergyPlus and CitySim models derived from a building information model, in: Proceedings of Bausim 2014: Fifth German-Austrian IBPSA Conference. Presented at the Bausim 2014: Fifth German-Austrian IBPSA Conference, IBPSA, Aachen, Germany, pp. 469–476. doi:10.13140/2.1.4639.1040\\[12pt]

Miller, C., Thomas, D., Kämpf, J., Schlueter, A., 2015. Long wave radiation exchange for urban scale modelling within a co-simulation environment, in: Proceedings of CISBAT 2015. Presented at the CISBAT 2015, CISBAT, Lausanne, Switzerland, pp. 871–876. doi:10.13140/RG.2.1.2876.4640\\[12pt]
This paper builds upon the previous research by applying the workflow to two campus case study projects. This contextual application gives a realistic analysis of how much differences between solo and co-simulations.

This paper was funded through the UMEM: Urban Multiscale Energy Modelling project from the Swiss Competence Center Energy and Mobility (CCEM).

Any questions or comments can be directed to myself as the corresponding author. I am avail-
able via email at miller.clayton@arch.ethz.ch and by phone at 1-402-403-0090.

\closing{Kind Regards,}

% Enclosures:\\
% Curriculum Vitae\\
% Reference List\\
% Research Statement\\
% Teaching Statement\\
% Unofficial Transcripts\\

%These insights will influence future design, construction, commissioning, and performance evaluation practices. 
%I believe Nebraska can become known for its building performance informatics research in addition to its already strong reputation for design and construction.

% It is through my active contacts and familiarity with Omaha and the State of Nebraska that I would be able to collect even larger data sets.
% more often applied to social media, advertising or medicine.

%I plan to undertake an academic career focused on analytics of performance data from the built and urban environments. My background includes experiences in engineering and architecture, as well as in research and entrepreneurship. In this application, I propose a teaching and research plan built upon this foundation.

%I thrive in interdisciplinary research environments. My primary advisor is an architect, my co-advisor is a computer science professor, my mentors are a robotics researcher and a mechanical engineer, and my colleagues are even more diverse. Crossing these boundaries is not trivial, and my career has a sound basis in this exact skill set.

%Sensor data are everywhere these days. As seen in the hype cycle, the Internet-of-Things concept is on many researchers' minds. Despite this fact, these ideas are not novel in the building industry; we have been connecting machines and sensors to the Internet for years. However, due to this movement's emphasis on sensor data, we rapidly realize how much more value can be extracted using new techniques. I am working at the forefront of utilizing temporal sensor data sets containing billions of measurements to improve performance, comfort and safety in buildings.

%HERE DISCUSS THE PROJECTS THAT I'M CURRENTLY COMPLETING AND THE FUNDING OPPORTUNITIES THAT I AM EXPLORING!

% It can be noticed from my CV that I am in the process of completing my Ph.D. degree. I have a deadline of June 2016 to submit my dissertation, and I plan to defend in the early Fall semester. This schedule puts me on track to be available for a faculty position in 2017. I see that the current on-line posting indicates a \emph{Planned Hire Date} of August 2016, but I'd like the opportunity to discuss further my application in the context of the Durham School's plans for expansion.

% \ps{Postscriptum}
% \encl{Research Statement\\
%  Teaching Statement}

% \cc{carbon copy}

% \newpage
% \setcounter{page}{1}
% \recipient{Address 2}
% \opening{Hallo all,}
% This is a sample for a serial letter. For the other changes please look into the file \emph{brief-sample-serial.tex}, there are only a few changes to the standard letter.

% In the moment there is only one known issue. In principal you can produce a serial letter with different pages each, at the moment I do not know, how the get in this circumstance the particular \emph{number of the last page}. This number could appear to be not the right one. The work-around is not to print the page-numbering (starting on the second page). Please uncomment in \emph{ethbrief3.cls} the following line then:

% \begin{verbatim}
% \hfill\raggedleft\mbox{\eth@xiiiv\ethpagename~\thepage~/~\lastpageref*{pagesLTS.arabic}}
% \end{verbatim}

% If anybody could help for this problem, I would like to include this in the following subversion.

% \closing{Have a nice day}
\end{document}